---
title: "Landsat 9 pixel cloud quality in R"
author: "Matthieu"
date: "March 8, 2019"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objective

This R script help *decode* the Landsat 8 `pixel_qa` band, and to recover the associated attributes. 

The Landsat 8 surface reflectance (SR) user manual [LANDSAT 8 SURFACE REFLECTANCE CODE (LASRC) PRODUCT GUIDE](https://www.google.com/url?sa=t&rct=j&q=&esrc=s&source=web&cd=1&cad=rja&uact=8&ved=2ahUKEwisnNbE5vPgAhVMvJ4KHUZwC0MQFjAAegQIChAC&url=https%3A%2F%2Flandsat.usgs.gov%2Fdocuments%2Flasrc_product_guide.pdf&usg=AOvVaw1k4ElRQCyGumQtZzeTT51P) is quite confusing, didn't find much helpful information there. Also, Landsat 8 stuff provides numerical tables only in... pdf format (confirmed by email to their helpline).

Some useful links were:

* [GIS Stackexchange: MOD09A1 QC layer](https://gis.stackexchange.com/questions/182924/mod09a1-qc-layer/252404#252404)
* [GIS Stackexchange: How can I parse modis MOD13Q1 quality layers in R?](https://gis.stackexchange.com/questions/144441/how-can-i-parse-modis-mod13q1-quality-layers-in-r?rq=1)
* [GIS Stackexchange: Landsat 8 cloud band: how to understand their (bit) coding?](https://gis.stackexchange.com/questions/314792/landsat-8-cloud-band-how-to-understand-their-bit-coding)

What I provide here:

* the function `lsqa_pixel_table()`, to extract cloud attributes from a Landsat 8 SR `pixel_qa` value
* two saved datasets in csv format, corresponding to [Table 7-3](https://raw.githubusercontent.com/MatthieuStigler/Misc/master/spatial/landsat_8_cloud/LS_8_pixel_qa_Table_7_3.csv) and [Table 7-4](https://raw.githubusercontent.com/MatthieuStigler/Misc/master/spatial/landsat_8_cloud/LS_8_pixel_qa_Table_7_4.csv) in the manual

# Usage: R script

Load the script hosted on github:

```{r, message = FALSE}
library(devtools)
source_url("https://raw.githubusercontent.com/MatthieuStigler/Misc/master/spatial/landsat_8_cloud/lsqa_tools.R")
```

## Get the attributes for one pixel value, say 322

```{r}
lsqa_pixel_table(322) %>% 
  spread(attribute, bit_value)
```


## Construct Table 7-4

This would be the list of all pixel values you can find in the data:

```{r}
pixel_qas <- c(1, 322, 324, 328, 336, 352, 368, 386, 388, 392, 400, 416, 432, 480, 834, 836, 840, 848, 864, 880, 898, 900, 904, 912, 928, 944, 
992, 1346, 1348, 1350, 1352)
```


Use now `lsqa_pixel_table()` and a few tidying steps:

```{r}
TAB_7_3 <- tibble(pixel_qa = pixel_qas) %>% 
  mutate(vals = map(pixel_qa, lsqa_pixel_table)) %>% 
  unnest(vals) %>% 
  spread(attribute, bit_value) %>% 
  mutate(pixel_qa = as.integer(pixel_qa)) %>% 
  arrange(pixel_qa) %>% 
  mutate_at(c("Fill", "Clear", "Water", "Cloud_Shadow", "Snow", "Cloud", "Terrain_Occlusion"), as.logical)

```

You could also get this table by downloading it directly:

```{r, include= FALSE}
read_csv("https://raw.githubusercontent.com/MatthieuStigler/Misc/master/spatial/landsat_8_cloud/LS_8_pixel_qa_Table_7_4.csv")
```

Results looks like: 

```{r, echo = FALSE}
TAB_7_3 %>% 
  knitr::kable()
```


## So when is it `Clear`?

Trying to understand how they classify their data... It looks like to be declared `Clear==TRUE` you need:

1. No Snow/Water (well actually yes: Water, but only if occluded?)
2. No Cloud_shadow or Cloud

You can have high cirrus, but not high Cloud?

```{r}
TAB_7_3 %>% 
  filter(Clear == TRUE) %>% 
  select(pixel_qa, Clear, Cloud, Water, Cloud_confidence, Cirrus_Confidence)
```

Check cases when  `Clear==FALSE`?

```{r}
TAB_7_3 %>% 
  filter(Clear == FALSE) 

```



 