---
title: 'Interpolating: alternative function'
author: "Matthieu"
date: "April 23, 2019"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R script to interpolate values ina  tidy way, reusing the distance matrix

This is a script I write some time ago. Objective was to extend `gstat::idw0()`:

- provide a function that recycles the distance matrix, hoping this would save tme
- allows for NAs in data (i.e. skip first neighbour if is NA)

# The function

The function: `idw_tidy()`. A bunch of arguments:

- data, newdata: as in `iwd()`
- idp = 2, maxdist=Inf, nmin=0, nmax=Inf: as in `iwd()`
- D=NULL, new!
- na.rm=TRUE, new!
- force=FALSE, parallel = NULL: new

# Demo

See the other file *idw_dplyr_USE.R* for a full demo.

Here we illustrate how to handle multiple variables:

```{r}
library(gstat)
library(sp)
library(sf)
library(dplyr)
library(purrr)

devtools::source_url("https://raw.githubusercontent.com/MatthieuStigler/Misc/master/spatial/inter_same_Dmat/idw_dplyr.R")


################################
#'## Load data
################################

data(meuse)
coordinates(meuse) = ~x+y
data(meuse.grid)
gridded(meuse.grid) = ~x+y
meuse_sf <- st_as_sf(meuse)
meuse.grid_sf <- st_as_sf(meuse.grid) |> select(geometry)


################################
#'## Simple use
################################

idw_tidy(data = meuse_sf  |> select(cadmium),
         newdata = meuse.grid_sf) |>
  head()


################################
#'## Use with many variables
################################


out_multi_tidy <- idw_tidy(data = meuse_sf  |> select(cadmium, copper, lead, zinc),
                      newdata = meuse.grid_sf) |>
  st_set_geometry(NULL)

head(out_multi_tidy)

### gstat way
get_dat <- function(var){
  out <- gstat(id = var, formula = as.formula(paste(var, "~ 1")), data = meuse) |> 
    predict(meuse.grid) 
  out2 <- out@data
  rownames(out2) <- NULL
  out2[,1,drop=FALSE]
}
  
out_multi_gstat <- purrr::map_dfc(c("cadmium", "copper", "lead", "zinc"),
    get_dat)

## Compare both ways
all.equal(out_multi_tidy, out_multi_gstat, check.attributes=FALSE)
```

