---
title: "README"
author: "Matthieu"
date: "2023-02-11"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code to get hansen data with rgee

This code simplifies the procedure to get Hansen forest loss data, including an initial step to mask pixels using the `treecover2000` raster layer. 

## Demo

Source script:

```{r}
devtools::source_url("https://raw.githubusercontent.com/MatthieuStigler/Misc/master/spatial/rgee/rgee_hansen_forest/eegfw_get_hansen.R")
```

Load libraries, authenticate:

```{r}
library(dplyr, warn.conflicts = FALSE)
library(sf)
library(rgee)

ee_Initialize(user = "XXX", gcs = FALSE)
## set to `gcs = TRUE` if you want to export after with ee_table_to_gcs()
```

Create pseudo input polygons:

```{r}
geom_1 <- ee$Geometry$Polygon(list(list(c(-73.83, 2.43) ,c(-73.83, 2.02),
                                        c(-73.14, 2.02),c(-73.14, 2.43))))
geom_2 <- ee$Geometry$Polygon(list(list(c(-73.72, 1.86) ,c(-73.72, 1.54),
                                        c(-72.88, 1.54), c(-72.88, 1.86))))
FC = ee$FeatureCollection(list(ee$Feature(geom_1, list(id="A")),
                               ee$Feature(geom_2, list(id="B"))))
ee_as_sf(FC)
```



### Run without mask:

```{r}
out_EE <- eegfw_get_dfrt(FC = FC)
GFW_masked_unmasked <- eegfw_quick_process(ee=out_EE)
GFW_masked_unmasked
```

## Run with mask

```{r}
mask_90 <- eegfw_get_mask()
out_mask_EE <- eegfw_get_dfrt(FC = FC, mask =mask_90)
GFW_masked <- eegfw_quick_process(ee=out_mask_EE)
GFW_masked
```


## Use TMF instead

To use the TMF product, use:

```{r}
TMF_v2022 <- eegfw_quick_process(ee=eeTMF_get_dfrt(FC = FC))
TMF_v2021 <- eegfw_quick_process(ee=eeTMF_get_dfrt(FC = FC, version='projects/JRC/TMF/v1_2021/DeforestationYear'))
```

## combine results and plot

```{r}
GFW_TMF <- rbind(TMF_v2022 %>% 
                   mutate(source= "TMF"),
                 GFW_masked %>% 
                   select(area_ee, id, DeforestationYear=lossyear, sum) %>% 
                   mutate(source = "GFW",
                          DeforestationYear=if_else(DeforestationYear==0, 0, DeforestationYear+2000)))

library(ggplot2)
GFW_TMF %>% 
  filter(DeforestationYear!=0) %>% 
  ggplot(aes(x=DeforestationYear, y=sum, color=id, linetype=source))+
  geom_line()+
  facet_wrap(~id, scales="free", ncol=1)+
  ggtitle("Comparing GFW and TMF (omitting 0 year = no def)")
```

## Export task

Eventually, export task using `rgee::ee_table_to_drive()` for google drive or `rgee::ee_table_to_gcs()` for google cloud storage. For the latter, make sure you initialized `ee_Initialize(user = "XXX", gcs = TRUE)` with `gcs = TRUE`. 

Example with `ee_table_to_gcs()`:

```{r, eval=FALSE}
## Try export
ee_table_to_gcs(
  collection=coll_out,
  description = desc,
  bucket = ,###,
  fileNamePrefix = name,
  timePrefix = FALSE,
  fileFormat = "geojson")
```

