---
title: "README"
output:
  html_document:
    keep_md: yes
date: "2023-01-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Voronoi tesselation between polygons

In R, `st_voronoi()` only works between points. To make it work between polygons, I implement a simple hack: convert polygons to point, then apply Voronoi, then merge polygons from same units. 

CAUTION: this DOES NOT WORK WELL when polygons are overlapping/touching! 

## Use

```{r cars}
source("stp_voronoi_polygons.R")
# source_url("https://raw.githubusercontent.com/MatthieuStigler/Misc/master/spatial/voronoi_polygons/stp_voronoi_polygons.R")
```

Create some data: countries in EU


```{r}
library(dplyr, warn.conflicts = FALSE)
library(sf)
library(rnaturalearth)

## Download and project
all_countries <- rnaturalearth::ne_countries(scale = "small", returnclass = 'sf')|> 
  dplyr::select(type, admin)  |> 
  st_transform("epsg:3035")

## filter to some in Europe
some_countries <- all_countries |> 
  dplyr::filter(admin %in% c("Italy", "Switzerland", "Ukraine", "Denmark", "Spain",
                             "Poland", "United Kingdom", "Norway", "Finland", "Greece"))
```

### Simple use

Works but not well when polygons touch, as for countries:

```{r}
some_countries_voro <- stp_voronoi_polygons(sf = some_countries, id_col = "admin") |> 
  st_intersection(st_bbox(some_countries) |> st_as_sfc())

col_alpha <- rgb(0, 0, 255, max = 255, alpha = 50)
plot(some_countries_voro |> st_geometry())
plot(some_countries |> st_geometry(), add=TRUE,col=col_alpha)
```

Result works in general, except for the polygons that are touching.

### Hack: slightly shrink countries (with `st_buffer()`) so that they don't touch

```{r}
## Shrink
some_countries_smaller <- some_countries %>% 
  st_buffer(-1000) %>% 
  nngeo::st_remove_holes() %>% 
  st_make_valid()

## Now apply
some_countries_smaller_voro <- stp_voronoi_polygons(sf = some_countries_smaller, id_col = "admin") |> 
  st_intersection(st_bbox(some_countries) |> st_as_sfc())

plot(some_countries_smaller_voro |> st_geometry())
```

### When output has spurious holes: add `buffer_inner`

Sometimes output will contain some holes, those could be produced by making the union of the pointwise-Voronoi polygons. TO handle this, add ing a very small buffer value should be good enough: 

```{r}
## sometimes need to add some small value before union
some_countries_smaller_voro_buf <- stp_voronoi_polygons(sf = some_countries_smaller,
                                                        id_col = "admin", buffer_inner = 1) |> 
  st_intersection(st_bbox(some_countries) |> st_as_sfc())

plot(some_countries_smaller_voro_buf |> st_geometry())
plot(some_countries_smaller |> st_geometry(), add=TRUE, col=col_alpha)
```

