---
title: "README"
author: "Matthieu"
date: "2023-02-19"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R script to check overlap of polygons

This script compute the dyadic (1-1) overlap of polygons among a given collection with `ovr_get_overlap_pairs`. It returns a data frame where each row is an overlapping dyad A-B, with respective areas and percentage of overlap with respect to A and B. 

Function `ovr_add_group()`  uses `igraph` tools to indicate groups of overlapping polygons int eh output from `ovr_get_overlap_pairs`. 

## Functions description:

- `ovr_get_overlap_pairs()`:
  - @param `sf` the sf object
  - @param `id_var` optional variable containing the row identifier
  - @param `unit` the unit in which to report the area
  - @param `pre_filter` whether to run first `st_intersects` to filter? Recommended as seems always
  faster...
  - @param `inter_make_valid` whether to repair potentially invalid values
  - @returns a tibble where each row represents an overlapping A-B dyad,  and columns indicate the id of the dyad, the area of each polygon, their interscetion and  percentage of overlap
- `ovr_add_group()`:
  - @param df_inter the output from `ovr_get_overlap_pairs`

## demo


Load code:
```{r}
source("https://raw.githubusercontent.com/MatthieuStigler/Misc/master/spatial/check_overlap/check_overlap.R")
```


Create demo data

```{r data}
library(sf)
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
  
M1 = matrix(c(0,0,10,0,10,10,0,10,0,0),ncol=2, byrow=TRUE)
M2 <- M1+0.01
M3 = matrix(c(1,1,1,4,4,4,4,1,1,1),ncol=2, byrow=TRUE)
M4 <- M3-5
M5 <- M3-6
M6 <- M3-7
M7 = matrix(c(11,11,11,12,12,12,12,11,11,11),ncol=2, byrow=TRUE)
M8 = matrix(c(0,-2,4,-2,4,-4,0,-4,0,-2),ncol=2, byrow=TRUE)
M9 <- M8+1

M_all <- list(M1, M2, M3, M4, M5, M6, M7, M8, M9)
pl2 = purrr::map(M_all, ~st_polygon(list(.)))
FC <- st_sf(ids = LETTERS[1:length(pl2)], geometry=st_sfc(pl2), x=rnorm(length(pl2)))
```

Plot the demo data:

```{r fig_data_raw}
ggplot(data = FC) +
  geom_sf(aes(fill = ids), alpha = 0.6) +
  geom_sf_text(aes(label = ids))
```

Use now the function:

```{r}
ovr_get_overlap_pairs(sf=FC)
```

Add groups:

```{r}
ovr_get_overlap_pairs(sf=FC) %>% 
  ovr_add_group()
```

Optional: specifiy polygons id variable

```{r}
ovr_get_overlap_pairs(sf=FC, id_var =ids)
```

Optional: run a first filter with `st_intersects`. In some cases where only a few features are overlapping, I found this to be faster. 

```{r}
all.equal(ovr_get_overlap_pairs(sf=FC, pre_filter = TRUE, id_var =ids) |> arrange(row_A, row_B),
          ovr_get_overlap_pairs(sf=FC, pre_filter = FALSE, id_var =ids)|> arrange(row_A, row_B))
```


