---
title: "README"
author: "Matthieu"
date: "2023-07-17"
output:
  html_document:
    toc: true
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction: Forest cover change scripts 

A collection of R scripts to process forest cover data from Hansen or PRODES.

Use with:

``` {r, eval=FALSE}
source("https://raw.githubusercontent.com/MatthieuStigler/Misc/master/spatial/forest_cover_change/frst_forest_cover_change_scripts.R")
```

# Output variables names

- `dfrt_year` year
- `dfrt_any` unit level: was there any observed deforestation for this unit over the whole sample?

Deforestation variables:

- `dfrt_area`: raw deforestation area
- `dfrt_area_by_tot`: deforestation area by unit area (%)
- `dfrt_area_by_forest`: deforestation area by initial forest area (%)
- `dfrt_area_annualized`: deforestation area by previous year's forest area (%)

Forest cover variables:

- `frst_area`: raw forest cover
- `frst_area_by_tot`: forest cover by unit area (%)


# Conventions:

When initial forest ==0:

 - the `dfrt_area_*` variables will be NA
 
When initial forest >0 but there is no deforestation:

- The output of `frst_HAN_process` will contain `dfrt_year` of `NA` (to make sure the observations appears once), but will have `dfrt_area_*` of 0
- The output of `frst_dfrt_complete` will contain `dfrt_year` with same years as other units
    
 

# Functions

- `frst_HAN_process`: main function
- `frst_dfrt_complete`: add years without deforestation
- `frst_HAN_add_frst_cover`: add forest cover


Check output:

- `frst_HAN_check_final`: main test
- `frst_add_forest_check`:


# Example

```{r, echo=FALSE}
suppressPackageStartupMessages(library(tidyverse, warn.conflicts = FALSE))
```

## Example 1: data does NOT contain unit's area

```{r}
library(tidyverse, warn.conflicts = FALSE)
source("https://raw.githubusercontent.com/MatthieuStigler/Misc/master/spatial/forest_cover_change/frst_forest_cover_change_scripts.R")

df_test_Han <- tibble(cell_id=rep(c("Normal", "No def", "Full def"), c(3,1, 1)),
                        lossyear = as.integer(c(0,2,3, 0, 5)),
                        area = 10000) %>%   
    group_by(cell_id) %>%
    mutate(area_total = sum(area)) %>%
    ungroup()
knitr::kable(df_test_Han)
```


```{r}
df_test_Han_prep <- df_test_Han %>% 
    frst_HAN_process(recompute_forest=TRUE) %>% 
    frst_dfrt_complete(nest_vars=nesting(cell_id, area_forest_initial, dfrt_any, area_total, area_forest_final),
                      years_all=2000:2005) 
df_test_Han_prep
```


## Example 2: data contains unit's area

```{r}
df_test_Han_2 <- tibble(cell_id=rep(c("Normal", "Zero def", "No deforestable", "some def"), c(3,1, 1,1)),
                          lossyear = as.integer(c(0,2,3,0,NA,5)),
                          area = 10000) %>%   
    group_by(cell_id) %>%
    mutate(area_total = sum(area)) %>%
    ungroup() %>% 
    mutate(area_forest_initial = if_else(is.na(lossyear), 0, area_total))

knitr::kable(df_test_Han_2)
```

```{r}
df_test_Han_prep_2 <- df_test_Han_2 %>% 
    frst_HAN_process(area_mask_forest_var = area_forest_initial) %>% 
    frst_dfrt_complete(nest_vars=nesting(cell_id, area_forest_initial, dfrt_any, area_total, area_forest_final),
                      years_all=2000:2005) 
df_test_Han_prep_2 %>% select(cell_id, area_total, area_forest_initial, dfrt_any, dfrt_year, dfrt_area)
```

## Example 3: Hansen data in the Amazonas

Data was process with rgee and is stored on:

```{r}
hansen_Amaz_demo <- read_csv("https://raw.githubusercontent.com/MatthieuStigler/Misc/master/spatial/forest_cover_change/hansen_raw_amazonas_BRA_from_rgee.csv",
                             show_col_types = FALSE)
```

Do full processing:

```{r}
hansen_Amaz_demo_clean <- hansen_Amaz_demo %>% 
  frst_HAN_process(area_mask_forest_var = mask_hansen, full_area_var = area_ee, area =sum,
                   .group_vars = c(ADM0_NAME, ADM1_NAME, ADM2_CODE, ADM2_NAME)) %>% 
  frst_dfrt_complete(nest_vars=nesting(ADM0_NAME, ADM1_NAME, ADM2_CODE, ADM2_NAME, 
                                       area_forest_initial, dfrt_any, area_total, area_forest_final)) %>% 
  frst_add_forest(.group_vars = ADM2_NAME, add_first_year = TRUE) %>% 
  frst_add_dfrt_annualized(.group_vars = ADM2_NAME)
```

Check the data, need to use a rather low tolerance to get all tests to pass:

```{r}
hansen_Amaz_demo_clean %>% 
  frst_HAN_check_final(.group_vars = ADM2_NAME, tol = 1e-6, has_forest_cover_0_year = TRUE)

```

Plot: turn to long and plot:

```{r, plot_Hansen_facets}
hansen_Amaz_demo_clean %>% 
  gather(dfrt_var, dfrt_value, starts_with(c("dfrt_area", "frst_area"))) %>% 
  ggplot(aes(x=dfrt_year, y= dfrt_value, group= ADM2_NAME))+
  geom_line(alpha=0.5)+
  facet_wrap(~dfrt_var, scales="free", ncol=2)+
  ggtitle("Municipalities in the Amazonas state, Brazil")
```

