---
title: "raster distances in R: different measures between cells"
author: "Matthieu"
date: "December 5, 2018"
output:
  md_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I wanted to check how to compute distances between cells ina  raster. There is the function `raster::distance()`, but it does not seem to be doing quite what I want. So I ended up using `rasterToPolygons()`, and either taking standard distances from the centroids, or using spatial distances found in package `spdep`. 

## Data setup

```{r, message = FALSE }
library(raster)
library(spdep)
library(tidyverse)
library(magrittr)
```


Make a very simple raster

```{r}
r <- raster(nrows = 9, ncols = 9, xmn=-1, xmx=1, ymn=-1, ymx=1)
# r[] <-  sample(1:81, size = 81, replace = FALSE)
```


### Extract cell coordinates:

```{r}
## Method 1
xyFromCell(r, cell = 1:ncell(r)) %>%  head
  

## Method 2
rasterToPolygons(r) %>%  
  coordinates %>% 
  head
```


### Distance using dist() on cells coordinates

```{r}
dist_rast <-  function(x, method = "euclidean")  {
  x %>% 
   xyFromCell(cell = 1:ncell(.)) %>% 
    dist(method = method) 
}
```


```{r}
r_new <- data_frame(dist = c("euclidean", "maximum", "manhattan", "canberra", "binary", "minkowski")) %>% 
  mutate(dist_dat =map(dist, ~ {r[] <- dist_rast(r, method = .) %>%  as.matrix() %>% .[41,]; names(r) <- .; r}) )  %$%
  stack(dist_dat)
plot(r_new)
```

Keep euclidean for later comparison:

```{r}
r$dist_euclid <- dist_rast(r) %>%  as.matrix() %>% .[41,]
```




### contiguous distance

```{r}
r$dist_contig <- rasterToPolygons(r)  %>% 
  poly2nb() %>% 
  nb2mat(style = "B") %>%  .[41,]
plot(r)
```


### contiguous: No queen

```{r}
r$dist_contig2 <- rasterToPolygons(r)  %>% 
  poly2nb(queen = FALSE) %>% 
  nb2mat(style = "B") %>%  .[41,]
plot(r$dist_contig2)
```

### Contiguous level 2

```{r}
r$dist_contig_rook_l2 <- rasterToPolygons(r)  %>% 
  poly2nb(queen = FALSE) %>% 
  nblag(maxlag=2) %>%  .[[2]] %>% 
  nb2mat(style = "B") %>%  . [41,]

r$dist_contig_queen_l2 <- rasterToPolygons(r)  %>% 
  poly2nb(queen = TRUE) %>% 
  nblag(maxlag=2) %>%  .[[2]] %>% 
  nb2mat(style = "B") %>%  . [41,]
```

```{r}
plot(r[[c("dist_contig_queen_l2", "dist_contig_rook_l2")]])
```

### Contiguous level 1 and 2

```{r}
cont_rook_lags <- rasterToPolygons(r)  %>% 
  poly2nb(queen = FALSE) %>% 
  nblag(maxlag=2) %>% 
  map(nb2mat, style = "B") 
cont_rook_l12 <- cont_rook_lags[[1]] + cont_rook_lags[[2]]*2
r$dist_contig_rook_l12 <- cont_rook_l12[41,]
plot(r$dist_contig_rook_l12)
```


```{r}
cont_queen_lags <- rasterToPolygons(r)  %>% 
  poly2nb(queen = TRUE) %>% 
  nblag(maxlag=2) %>% 
  map(nb2mat, style = "B") %>%  
  map2(list(1, 2), ~.x *.y ) %>% 
  {Reduce('+', .)}


r$dist_contig_queen_l12 <- cont_queen_lags[41,]
plot(r$dist_contig_queen_l12)
```

## Show all

```{r}
plot(r)
```

