---
title: "idw interpolation using spatially weighted regression"
author: "Matthieu"
date: "July 16, 2018"
output:
  md_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A common technique to do interpolation is to use inverse-distance weights to compute local means on new points. This is implemented in R for example in the package `gstat`, with funciton `idw0` or `gstat`. 

An equivalent way to do this is to use spatially weighted local regression (GWR). Remembering that regressing a variable against the intercept only will give the mean of the variable, same logic predicts that using a GWR model should provide a way to do interpolation. Package `spgwr` provides funciton `gwr` which takes either a Gaussian or bi-square weighting scheme. Implementing a idw weigthing scheme is fairly easy, yet results differ importantly so far. 

Three points that are unclear so far:

* `spgwr` versus `gstat`: why are `gwr()` and `gstat()` not giving identical results, Inifinite or restricted bandwidth ? **partial answer**: distances in `spgwr` are already squared, so to use idw at power 2 needs just: `1/dist2`
* gstat: why are `idw0()` and `gstat()` results not identical when `maxdist=Inf`?
* `spgwr` why are so many points receving NA? I use a bandwidth of 2, and noting that the maximim minimal distance is 1.07, every new-point should at least have one neighbour!?
 
## Loading and preparing data

Use Columbus data, split at random for a train and test sub-samples. 

```{r cars, message=FALSE}
library(tidyverse)
library(spgwr)
library(gstat)
library(rgdal)
library(knitr)

# columbus_sp <- readOGR(system.file("shapes/columbus.shp", package="spData")[1])
data(columbus, package="spData")

columbus_sp <- columbus
coordinates(columbus_sp) <- columbus[,c("X", "Y")]

set.seed(123)
test_ids <- sample(1:nrow(columbus), size=10)

col_XY <- coordinates(columbus_sp)

col_sp_train <- columbus_sp[-test_ids,]
col_sp_test <- columbus_sp[test_ids,]

col_train <- columbus[-test_ids,]
col_test <- columbus[test_ids,]
```



## Define the idw kernel for gwr

```{r}
## kernels
gwr_iwd <-  function (dist2, d)    {
  d2 <- d^2
  w <- ifelse(dist2 > d2, 0, 1/dist2)
  w
}
```


## Gwr and idw interpolation

### gwr ones: 

```{r}
### Regressions
gwr_gauss <- gwr(CRIME ~ 1, data=col_train,
                 fit.points = col_XY[test_ids,],
                 coords=col_XY[-test_ids,], bandwidth=2,
                 gweight=gwr.bisquare)
gwr_gauss$SDF$"(Intercept)"

gwr_gauss$SDF$"(Intercept)"


gwr_idw_2 <- gwr(CRIME ~ 1, data=col_train,
                 fit.points = col_XY[test_ids,],
                 coords=col_XY[-test_ids,], bandwidth=4,
                 gweight=gwr_iwd)

gwr_idw_Inf <- gwr(CRIME ~ 1, data=col_train,
                   fit.points = col_XY[test_ids,],
                   coords=col_XY[-test_ids,], bandwidth=Inf,
                   gweight=gwr_iwd)

gwr_idw_2$SDF$"(Intercept)"
```
### knn-style with gwr

For a knn, we play with the argument `adapt`. Threoteically, setting 2/39 should give us a 2knn, but it seems it give just above, so I set 2/39-0.001. Here I try knn wihtout idw weighting:

```{r}
gwr_knn <-  function (dist2, d)    {
  d2 <- d^2
  w <- ifelse(dist2 > d2, 0, 1)
  w
}
```

```{r}
gwr_knn2 <- gwr(CRIME ~ 1, data=col_train,
                fit.points = col_XY[test_ids,], adapt= 2/39,
                coords=col_XY[-test_ids,], bandwidth=Inf,
                gweight=gwr_knn)
as.data.frame(gwr_knn2$SDF)
```



### gstat ones:

```{r}
gstat_2_mod <- gstat(id = "CRIME", formula = CRIME ~ 1, data = columbus_sp[-test_ids,], 
                     maxdist = 4, set = list(idp = 2)) 
gstat_2 <- as.data.frame(predict(gstat_2_mod, columbus_sp[test_ids,]))
  
gstat_knn2_mod <- gstat(id = "CRIME", formula = CRIME ~ 1, data = columbus_sp[-test_ids,], 
                        nmax = 2, set = list(idp = 0)) 
gstat_knn2 <- as.data.frame(predict(gstat_knn2_mod, columbus_sp[test_ids,]))


gstat_Inf_mod <- gstat(id = "CRIME", formula = CRIME ~ 1, data = columbus_sp[-test_ids,], 
                 maxdist = Inf, set = list(idp = 2)) 
gstat_Inf <- as.data.frame(predict(gstat_Inf_mod, columbus_sp[test_ids,]))


##
idw_Inf <- idw0(CRIME ~ 1, data=columbus_sp[-test_ids,], newdata=columbus_sp[test_ids,])
```

## Compare results

```{r}
table <- data.frame(idw_Inf = as.numeric(idw_Inf),
                    gstat_Inf = gstat_Inf$CRIME.pred,
                    gwr_idw_Inf = gwr_idw_Inf$SDF$"(Intercept)",
                    # gwr_idw_Inf_old = gwr_idw_Inf_old$SDF$"(Intercept)",
                    gstat_bw = gstat_2$CRIME.pred,
                    gwr_idw_bw = gwr_idw_2$SDF$"(Intercept)",
                    # gwr_gauss = gwr_gauss$SDF$"(Intercept)",
                    gstat_knn2 = gstat_knn2$CRIME.pred,
                    gwr_knn2 = gwr_knn2$SDF$"(Intercept)")
kable(table, digits=2)
```

Summary of diffs:

```{r}
diff_sumry <- table %>%
  select(-idw_Inf) %>%
  mutate(n_row = 1:n()) %>%
  gather(variable, value, -n_row) %>%
  mutate(type= str_extract(variable, "gstat|gwr"),
         spec = str_extract(variable, "bw|Inf|knn")) %>%
  select(-variable) %>%
  spread(type, value) %>%
  mutate(diff = gstat-gwr) %>%
  group_by(spec) %>%
  summarise_at("diff", funs(min, max, mean))
kable(diff_sumry)
```


## Compute max-min distance



```{r}
D <- spDists(col_XY[test_ids,],
             col_XY[-test_ids,])

D <- spDists(columbus_sp[test_ids,],
             columbus_sp[-test_ids,])
## min distance from train point to each new/test point
D_min_each <- apply(D, 1, min, na.rm=TRUE)
sort(D_min_each)
```

#### Max of the min distance
```{r}
max(D_min_each)
```

#### compute sum of weights, and how many weights not zero

```{r}
apply(D, 1, function(x) sum(gwr_iwd(x^2, 2)))
apply(D, 1, function(x) sum(gwr_iwd(x^2, 2)>0))
```

#### Find out the distance used in grw:

If we get use knn-1 and return as weight the (non-inverse) distance, we should find out which distances are used in grw:

```{r}
as.data.frame(gwr(CRIME ~ 1, data=col_train,
    fit.points = col_XY[test_ids,],
    coords=col_XY[-test_ids,], adapt=1/39-0.000001,
    gweight=function(dist2, d) ifelse(dist2>d^2, 0, sqrt(dist2)))$SDF)

```

