---
title: "idw interpolation using spatially weighted regression"
author: "Matthieu"
date: "July 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A common technique to do interpolation is to use inverse-distance weights to compute local means on new points. This is implemented in R for example in the package `gstat`, with funciton `idw0` or `gstat`. 

An equivalent way to do this is to use spatially weighted local regression (GWR). Remembering that regressing a variable against the intercept only will give the mean of the variable, same logic predicts that using a GWR model should provide a way to do interpolation. Package `spgwr` provides funciton `gwr` which takes either a Gaussian or bi-square weighting scheme. Implementing a idw weigthing scheme is fairly easy, yet results differ importantly so far. 

Three points that are unclear so far:

* `spgwr` versus `gstat`: why are `gwr()` and `gstat()` not giving identical results, Inifinite or restricted bandwidth ?
* gstat: why are `idw0()` and `gstat()` results no identical when `maxdist=Inf`?
* `spgwr` why are so many points receving NA? I use a bandwidth of 2, and noting that the maximim minimal distance is 1.07, every new-point should at least have one neighbour!?
 
## Loading and preparing data

Use Columbus data, split at random for a train and test sub-samples. 

```{r cars, message=FALSE}
library(tidyverse)
library(spgwr)
library(gstat)
library(rgdal)

columbus_sp <- readOGR(system.file("shapes/columbus.shp", package="spData")[1])
data(columbus, package="spData")


set.seed(123)
test_ids <- sample(1:nrow(columbus), size=10)

col_XY <- as.matrix(columbus[,c("X", "Y")])

col_sp_train <- columbus_sp[-test_ids,]
col_sp_test <- columbus_sp[test_ids,]

col_train <- columbus[-test_ids,]
col_test <- columbus[test_ids,]
```

## Define the idw kernel for gwr

```{r}
## kernels
gwr_iwd <-  function (dist2, d, pow=2)    {
  w <- ifelse(dist2 > d, 0, 1/(dist2^pow))
  w
}
```


## Gwr and idw interpolation

### gwr ones: 

```{r}
### Regressions
gwr_gauss <- gwr(CRIME ~ 1, data=col_train,
           fit.points = col_XY[test_ids,],
           coords=col_XY[-test_ids,], bandwidth=2,
           gweight=gwr.bisquare)
gwr_gauss$SDF$"(Intercept)"

gwr_gauss$SDF$"(Intercept)"


gwr_idw_2 <- gwr(CRIME ~ 1, data=col_train,
                fit.points = col_XY[test_ids,],
                coords=col_XY[-test_ids,], bandwidth=2,
                gweight=gwr_iwd)

gwr_idw_Inf <- gwr(CRIME ~ 1, data=col_train,
                 fit.points = col_XY[test_ids,],
                 coords=col_XY[-test_ids,], bandwidth=Inf,
                 gweight=gwr_iwd)
gwr_idw_2$SDF$"(Intercept)"
```

#### gstat ones:

```{r}
gstat_2 <- gstat(id = "CRIME", formula = CRIME ~ 1, data = columbus_sp[-test_ids,], 
              maxdist = 2, set = list(idp = 2)) %>%
  predict(columbus_sp[test_ids,]) %>%
  as_data_frame()

gstat_Inf <- gstat(id = "CRIME", formula = CRIME ~ 1, data = columbus_sp[-test_ids,], 
                 maxdist = Inf, set = list(idp = 2)) %>%
  predict(columbus_sp[test_ids,]) %>%
  as_data_frame()

idw_Inf <- idw0(CRIME ~ 1, data=columbus[-test_ids,], newdata=columbus[test_ids,]) # ,y=columbus[-test_ids,"CRIME"]
```

## Compare results

```{r}
table <- data_frame(idw_Inf = as.numeric(idw_Inf),
                   gstat_Inf = gstat_Inf$CRIME.pred,
                   gstat_2 = gstat_2$CRIME.pred,
                   gwr_idw_Inf = gwr_idw_Inf$SDF$"(Intercept)",
                   gwr_gauss = gwr_gauss$SDF$"(Intercept)",
                   gwr_idw_2 = gwr_idw_2$SDF$"(Intercept)")
table
```

```{r}
D <- spDists(coordinates(columbus_sp[test_ids,]),
             coordinates(columbus_sp[test_ids,]))
diag(D) <- NA
apply(D, 1, min, na.rm=TRUE)
max(apply(D, 1, min, na.rm=TRUE))
```
