---
title: "Spatial custering with ClustGeo"
author: "Matthieu"
date: "December 6, 2018"
output:
  md_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(raster)
library(tidyverse)
library(ClustGeo)
library(cowplot)
```


## Intro

I want to identify parts in a field with lower yields, and want these parts to be contiguous. In spatial word, I am trying to do a 1D clustering, taking into account the spatial distribution of the data. I use here  R package `ClustGeo` provides function `hclustgeo()` that does a mixing of data and spatial distance doing *Ward-like hierarchical clustering algorithm including soft contiguity constraints.*

`ClustGeo` has a parameter $\alpha$, that indicates how much we want to emphasize *data-clustering*  versus *spatial-clustering* 

* $\alpha = 0$: only data clustering
* $\alpha > 0$ also give weight to saptial-clustering

Here, I give an example with a field with **random values** (zero spatial correlation). Under random (non-spatial) sampling I expect:

* data: with increasing $\alpha$, means of cluster should become less and less different
* spatial: with increasing $\alpha$, cluster should get closer to...  a square probably (that's the form with the lowest average difference, correct?) 



## Functions: prep data, compute clustering for various alpha

### Pseudo data



```{r pressure}
r <- raster(nrows = 9, ncols = 9, xmn=-1, xmx=1, ymn=-1, ymx=1)
r[] <-  sample(1:81, size = 81, replace = FALSE)
plot(r)
```

Convert to simple data

```{r}
ras_dat <- as.data.frame(r, xy = TRUE) %>%  as_tibble %>% 
  mutate(n_cell = 1:ncell(r)) %>% 
  select(n_cell, everything()) %>% 
  rename(value = layer)
ras_dat
```

### Warm-up: some convenience functions

Just make a distance function, and a nicer (data_frame) output for the `hclustgeo()` function. 

```{r}
dist_rast_euclid <-  function(x)  {
  x %>% 
    xyFromCell(cell = 1:ncell(.))  %>% 
    dist() 
}
hclustgeo_df <-  function(D0, D1 = NULL, alpha, n_obs = TRUE, k = 5) {
  res <- hclustgeo(D0, D1, alpha = alpha) %>% 
    cutree(k=k) %>% 
    data_frame(cluster = .)
  if(n_obs) res <-  res %>% 
      mutate(n_obs =   1:nrow(.)) %>% 
      select(n_obs, everything())
  res
  
}
```



### Compute distances

```{r}
dat_dist <- dist(getValues(r$layer))
geo_dist <-  dist_rast_euclid(r)
```


### Use function on grid of alphas

I use alphas from 0 to 1, by 0.1.

```{r}
res_alphas <- data_frame(alpha = seq(0, 1, by = 0.1)) %>% 
  mutate(alpha_name = paste("alpha", alpha, sep="_"),
         data = map(alpha, ~ hclustgeo_df(dat_dist, geo_dist, alpha = ., k=5)))

res_alphas_l <-  res_alphas %>% 
  unnest(data) %>% 
  left_join(ras_dat, by = c("n_obs" = "n_cell")) %>% 
  mutate_at(c("alpha", "cluster"), as.factor) %>% 
  group_by(alpha, cluster) %>% 
  mutate(cluster_mean = mean(value)) %>% 
  ungroup()
```



## Result: Visualisation

```{r}
res_alphas_l_dat <-  res_alphas_l %>% 
  filter(alpha %in% c(0, 0.1, 0.5, 0.8, 0.9, 1)) 
```

### Various plots

```{r}
## show original
pl_dat_orig <- res_alphas_l_dat %>% 
  filter(alpha ==0) %>% 
  ggplot(aes(x = x, y= y, fill = value)) +
  geom_tile() +
  ggtitle("Original data (no spatial corr actually)")
```

```{r}
## show clustering in dat space
pl_clus_datSpace <- res_alphas_l_dat %>% 
  ggplot(aes(x = value, y= cluster, colour = cluster)) +
  geom_point() +
  facet_grid(alpha ~ .) +
  theme(legend.position = "none") +
  ggtitle("Cluster, in data space")

pl_clus_datSpace

## show clustering in geo space
pl_clus_geoSpace <- res_alphas_l_dat %>% 
  ggplot(aes(x = x, y= y, fill = factor(cluster))) +
  geom_tile() +
  facet_grid(alpha ~ .) +
  theme(legend.position = "none") +
  ggtitle("Cluster, in geo-space")

pl_clus_geoSpace
```



#### Show means

```{r}
pl_clus_geoSpace_means <- res_alphas_l_dat %>% 
  ggplot(aes(x = x, y= y, fill = cluster_mean)) +
  geom_tile() +
  facet_grid(alpha ~ .) +
  theme(legend.position = "none") +
  ggtitle("Cluster, in geo-space, colour according to mean values")

pl_clus_geoSpace_means

```

yes indeed, as I expected, spatially-clustering random points reduces variability between cluster means!

### Final result: compare data- versus spatial-clustering

PLug together in one (pkg `cowplot`):

```{r}
plot_grid(pl_dat_orig, plot_grid(pl_clus_datSpace, pl_clus_geoSpace), nrow = 2, rel_heights = c(1, 2))
```

